Quantum H theorem, variational entropy, and relaxation processes

### 2021-01-11 tex/introduction.tex

In the literature, the theoretical basis that allows us to describe equilibrium
systems in the classical scheme is well-established.
The procedures to describe
equilibrium systems are well-known and work in almost every system in nature
(Thermodynamics for a phenomenological description and statistical mechanics
for a microscopic description). In contrast, a unified mathematical model to
describe the evolution of any system to an equilibrium state has not developed.
Nevertheless, developments can be found in the literature and describe the
evolution to equilibrium to almost any system (the internal behavior of glasses
\cite{bib:zanotto2018} and the entropy measurement
\cite{bib:schmelzer2018,bib:nemilov2018} are some examples where the recent
developments do not work.) Examples of those developments can be the Onsager
formulation (called linear thermodynamics)
\cite{bib:keizer1987,bib:onsager1931}, and the Kinetic Theory of dilute gases
developed by Boltzmann.

In the Kinetic Theory of dilute gases, the behavior of a dilute classical gas
was described by Boltzmann through the Boltzmann transport equation.
Furthermore, the evolution of an equilibrium state for any system was proved by
means the Boltzmann's $H$-theorem (or classical $H$-theorem) and the
spatial-homogeneous distribution function hypothesis.
However, there are some aspects in the classical $H$-theorem and the
Boltzmann's $H$-functional that deserve some review to accomplish better
consistency, for example, the modification of the $H$-theorem in stochastic
trajectories, violations in the second law of thermodynamics, and the relation
between Shannon's measure of information and the Boltzmann's entropy %[6-10].
\cite{bib:nemilov2018,bib:keizer1987,bib:onsager1931,bib:brown2008,bib:dragoljub2009}.

>>>>

In the literature, the theoretical basis that allows us to describe equilibrium
systems in the classical scheme are well-established.
The procedures to describe
equilibrium systems are well-known and work properly for the treatment of almost every system in nature
(Thermodynamics for a phenomenological description and statistical mechanics
for a microscopic description). In contrast, a unified mathematical model to
describe the evolution of any system from an arbitrary out of equilibrium condition to an equilibrium state has not been developed yet.
Nevertheless, developments can be found in the literature and describe the
evolution to equilibrium to almost any system (the internal behavior of glasses
\cite{bib:zanotto2018} and the entropy measurement
\cite{bib:schmelzer2018,bib:nemilov2018} are some examples where the recent
developments do not work.) Examples of those developments can be the Onsager
formulation (called linear thermodynamics)
\cite{bib:keizer1987,bib:onsager1931}, and the Kinetic Theory of dilute gases
developed by Boltzmann.

In the Kinetic Theory of gases, the behavior of a dilute classical gas
was described by Boltzmann through the Boltzmann transport equation.
Furthermore, the evolution of an equilibrium state for any system was proved by
means the Boltzmann's $H$-theorem (or classical $H$-theorem) and the
spatial-homogeneous distribution function hypothesis.
However, there are some aspects in the classical $H$-theorem and the
Boltzmann's $H$-functional that deserve some review to accomplish better
consistency, for example, the modification of the $H$-theorem in stochastic
trajectories, violations in the second law of thermodynamics, and the relation
between Shannon's measure of information and the Boltzmann's entropy %[6-10].
\cite{bib:nemilov2018,bib:keizer1987,bib:onsager1931,bib:brown2008,bib:dragoljub2009}.

### 2021-01-11 tex/quantum.tex

Assuming that the energy exchange between particles is produced by collisions between particles.
By using perturbation theory, the rate of change in the number of particles in a group $\kappa$ is
%
\begin{eqnarray}\label{eq:changen}
    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}).
\end{eqnarray}
%
Here $A_{\kappa\lambda,\mu\nu}n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})$ is the
expected number of collisions per unit time, in which two particles will be moved from groups
$\kappa$, $\lambda$ to $\mu$, $\nu$, and the tensor $A_{\kappa\lambda,\mu\nu}n_{\kappa}$ is given by
%
\begin{equation}\label{eq:amnkldef}
  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon}.
\end{equation}
%
In Eq.~(\ref{eq:amnkldef}), $\Delta\epsilon$ is the net energy change occuring during the
collision and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where
$V_{mn,kl}$ is the element of the transition matrix of a binary collision.
It is important to remark that in deriving Eq.~(\ref{eq:changen}), the equal a priori
probabilities and the random a priori phases hypotheses were taken valid. The 
the random a priori phases hypothesis can be
considered an analogous to the molecular chaos hypothesis
\cite{bib:das2018} as the mechanism to introduce stochasticity into the system.

>>>>

Assuming that the energy exchange between particles is produced by collisions between particles.
By using perturbation theory, the rate of change in the number of particles in a group $\kappa$ is
%
\begin{eqnarray}\label{eq:changen}
    \frac{d n_{\kappa}}{dt}&=&-\sum_{\lambda,(\mu \nu)}A_{\kappa\lambda,\mu\nu} n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})\nonumber \\
    &&+\sum_{\lambda,(\mu \nu)}A_{\mu\nu,\kappa\lambda} n_{\mu}n_{\nu}(g_{\kappa}\pm n_{\kappa})(g_{\lambda}\pm n_{\lambda}).
\end{eqnarray}
%
Here $A_{\kappa\lambda,\mu\nu}n_{\kappa}n_{\lambda}(g_{\mu}\pm n_{\mu})(g_{\nu}\pm n_{\nu})$ is the
expected number of collisions per unit time, in which two particles will be moved from groups
($\kappa$, $\lambda$) to ($\mu$, $\nu$), and the tensor $A_{\kappa\lambda,\mu\nu}$ is given by
%
\begin{equation}\label{eq:amnkldef}
  A_{\kappa\lambda,\mu\nu}=\frac{4\pi^{2}}{h}\frac{|I_1\pm I_2|^2}{\Delta \epsilon}.
\end{equation}
%
In Eq.~(\ref{eq:amnkldef}), $\Delta\epsilon$ is the net energy change occuring during the
collision and $|I_1-I_2|^2=|V_{mn,kl}|^2$ where
$V_{mn,kl}$ is the element of the transition matrix of a binary collision.
It is important to remark that in deriving Eq.~(\ref{eq:changen}), the equal a priori
probabilities and the random a priori phases hypotheses were taken valid. The 
the random a priori phases hypothesis can be
considered an analogous to the molecular chaos hypothesis
\cite{bib:das2018} as the mechanism to introduce stochasticity into the system.
